---
title: "Impact of noise on HMM estimations"
author: "Laura Symul"
date: "8/8/2018"
output: html_document
---

```{r impact_of_noise setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r impact_of_noise sourcing functions, variables and libraries, include=FALSE}
source("Scripts/_setup_Sympto.R")

theme_set(theme_light())
```


# Loading & selecting data


```{r impact_of_noise loading Sympto data, cache = TRUE}


load(file = paste0(IO$sympto_data_02_standard_cycles,'users.Rdata'))
load(file = paste0(IO$sympto_data_02_standard_cycles,'cycles.Rdata'))
load(file = paste0(IO$sympto_data_02_standard_cycles,'cycledays.Rdata'))

head(cycles)

```


```{r impact_of_noise selecting X cycles, cache = TRUE}

X = 100

cycle_ids = unique(cycles$cycle_id)

selected_cycle_ids = cycle_ids
selected_cycle_ids = cycle_ids[1:X]

```


```{r impact_of_noise showing the transition probabilities, cache= TRUE}

hmm_par$cycle.states.transProbs

```




# Running the original HMM (= the HMM of the paper so far)

```{r impact_of_noise running original HMM, cache = TRUE}

#### HMM
RES = data.frame()

tic()
for (cycle_id in selected_cycle_ids) {
  #cat(cycle_id, '\n')
  j = which(cycledays$cycle_id == cycle_id)
  cycletable = cycledays[j, ]
  
  res = try(most_likely_day_of_ovulation_hmm(cycletable = cycletable,
                                         debug = FALSE, no.print = TRUE),
            silent = TRUE)
  
  if(typeof(res)!= 'character'){
    res$cycle_id = cycle_id
    line = data.frame(
      cycle_id = cycle_id, 
      ovu = res$ovu, 
      ovu.sd = res$ovu.sd,
      ovu.viterbi = res$ovu.viterbi, 
      confidence = res$confidence,
      sequence = paste(res$obs.viterbi, collapse = " "))
    RES = rbind(RES, line)
    }
  
}
toc()
head(RES)
RES.o = RES

```



# Modifying the HMM transition probabilities

For each cycle, we run Y different HMM with noise on the transistion probabilities


```{r  impact_of_noise running modified HMM with changing transition probabilities, cache = TRUE}

Y = 9

RES = data.frame()

tic()
for (cycle_id in selected_cycle_ids) {
  #cat(cycle_id, '\n')
  j = which(cycledays$cycle_id == cycle_id)
  cycletable = cycledays[j, ]
  
  for(y in 1:Y){
    #cat("\t", y, "\n")
    
    #noise = -log10(y/Y*10)+1
    noise = y/Y *3
    
    res = try(most_likely_day_of_ovulation_hmm(cycletable = cycletable,
                                               debug = TRUE, no.print = TRUE,
                                               random_transition_probabilities = TRUE,
                                               noise = noise),
              silent = TRUE)
    
    if(typeof(res)!= 'character'){
      #cat("\t\there\n")
      res$cycle_id = cycle_id
      line = data.frame(
        cycle_id = cycle_id, 
        y = y,
        noise = noise,
        dev.sym = res$dev.sym,
        ovu = res$ovu, 
        ovu.sd = res$ovu.sd,
        ovu.viterbi = res$ovu.viterbi, 
        confidence = res$confidence,
        sequence = paste(res$obs.viterbi, collapse = " "))
      RES = rbind(RES, line)
    }
    
  }
  
}
toc()

m = match(RES$cycle_id, RES.o$cycle_id)
RES$ovu.no.noise = RES.o$ovu[m]
RES$ovu.diff = RES$ovu.no.noise - RES$ovu
RES$ovu.sd.no.noise = RES.o$ovu.sd[m]
RES$sequence.no.noise = RES.o$sequence[m]
RES$noise.r = round(RES$noise, digits = 2)

head(RES, 20)

```

```{r impact_of_noise  visualisation params, cache = TRUE}

save(RES, file = paste0(IO$output_Rdata, "RES_1.Rdata"))

```




```{r impact_of_noise visualizing results of noisy HMM, cache = TRUE}

xlim.max = 20

g.noise = ggplot(RES, aes(x = ovu.diff, fill = noise))
g.noise = g.noise + geom_histogram(bins = 100) + 
  facet_grid(noise.r ~.) + 
  scale_fill_gradient(low = "gray30", high = "tomato")+
  guides(fill=FALSE)+  
  scale_y_continuous(breaks = seq(0,100,by =30))+
  theme(strip.text.y = element_text(angle = 0))+
  #ggtitle("difference in ovulation estimation") + 
  xlab("Diff. in ovu. est. (no noise - noise) [days]") + xlim(-xlim.max,xlim.max) 

g.noise



g = ggplot(RES, aes(x = dev.sym, fill = factor(noise)))
g + geom_histogram(bins = 50)  + facet_grid(noise~.)


g.noise.scatter = ggplot(RES, aes(x = abs(ovu.diff), y = ovu.sd.no.noise, col = noise))
g.noise.scatter = g.noise.scatter  + 
  geom_abline(slope = 1, intercept = 0, col = "gray")+ 
  geom_point(size = 0.5) + 
  scale_color_gradient(low = "gray30", high = "tomato")+
  facet_wrap(noise.r ~., ncol = 5)+
  xlab("Abs. diff. in ovu. est. (no noise - noise) [days]")+
  ylab("Uncertainty on ovulation est. (no noise)")+
  guides(col=FALSE)  + xlim(0,xlim.max) 

g.noise.scatter

```


## Conclusion:

Even large variations in transition probabilities don't have a big impact on the ovulation day estimation. Also, the difference in ovulation day estimation is smaller than the estimated uncertainty when the noise is reasonable (= the transition probabilities are not completely random).




# What about the state sequence for the other states (not just ovulation)?

```{r impact_of_noise checking the state sequence differences, cache = TRUE}

STD = data.frame()

#plot(c(0,50),c(0,nrow(RES)), type = "n")
for(i in 1:nrow(RES)){
  seq = RES$sequence[i]
  seq = unlist(strsplit(as.character(seq), split = " "))
  seq.num = match(seq, hmm_par$cycle.states)
  
  seq.o = RES$sequence.no.noise[i]
  seq.o = unlist(strsplit(as.character(seq.o), split = " "))
  seq.o.num = match(seq.o, hmm_par$cycle.states)
  
  #points(rep(i,length(seq)), col = cycle.states.colors[seq.num], pch = "-")
  
  dist.seq = sum(seq.num != seq.o.num)
  n.back.trans = sum(diff(seq.num) == -1)
  period.duration = sum(seq.num %in% c(1,2))
  period.duration.o = sum(seq.o.num %in% c(1,2))
  
  line = data.frame(cycle_id = RES$cycle_id[i],
                    y = RES$y[i],
                    noise = RES$noise[i],
                    dist.seq = dist.seq,
                    cycle_length = length(seq.num),
                    n.back.trans = n.back.trans,
                    period.duration = period.duration,
                    period.duration.change = period.duration-period.duration.o)
  STD = rbind(STD, line)
  
}

STD$dist.seq.f = STD$dist.seq/STD$cycle_length * 100


g = ggplot(STD, aes(x = dist.seq.f, fill = factor(noise)))
g + geom_histogram() + facet_grid(noise~.)


g = ggplot(STD, aes(x = n.back.trans, fill = factor(noise)))
g + geom_histogram() + facet_grid(noise ~.)


g = ggplot(STD, aes(x = period.duration.change, fill = factor(noise)))
g + geom_histogram() + facet_grid(noise ~.)

```

## Conclusion

Non-uniform (random) transition probabilities have an impact on the state sequence (probably a larger impact on the state sequence than on the estimation of the ovulation day). Especially, we can see that it can lead to a lot more back & forth transitions between the substates of the ovulatory phase. 
It does not affect too much the period duration estimation (>> the observations are still leading the state transition when decoding).

It could be interesting to check how the more "conservative" transition probabilities (e.g. the probability to stay in a state is higher than the probability to shift to another state) would impact the sequence.


# More conservative transition probabilities

Instead of just randomizing transition probabilities, we could check how increasing the probabilities to stay in a given state


```{r impact_of_noise running modified HMM conservative transition probabilities, cache = TRUE}

Y = 10

RES.c = data.frame()
#trans.probs = data.frame()

tic()
for (cycle_id in selected_cycle_ids) {
  #cat(cycle_id, '\n')
  j = which(cycledays$cycle_id == cycle_id)
  cycletable = cycledays[j, ]
  
  for(y in 1:Y){
    #cat("\t", y, "\n")
    
    C = y
    
    res = try(most_likely_day_of_ovulation_hmm(cycletable = cycletable,
                                               debug = FALSE, no.print = TRUE,
                                               conservative_transition_probabilities = TRUE,
                                               C = C),
              silent = TRUE)
    
    
    if(typeof(res)!= 'character'){
      #cat("\t\there\n")
      res$cycle_id = cycle_id
      line = data.frame(
        cycle_id = cycle_id, 
        y = y,
        C = C,
        dev.sym = res$dev.sym,
        ovu = res$ovu, 
        ovu.sd = res$ovu.sd,
        ovu.viterbi = res$ovu.viterbi, 
        confidence = res$confidence,
        sequence = paste(res$obs.viterbi, collapse = " "))
      RES.c = rbind(RES.c, line)
    }
    
  }
  
}
toc()


m = match(RES.c$cycle_id, RES.o$cycle_id)
RES.c$ovu.no.noise = RES.o$ovu[m]
RES.c$ovu.diff = RES.c$ovu.no.noise - RES.c$ovu
RES.c$ovu.sd.no.noise = RES.o$ovu.sd[m]
RES.c$sequence.no.noise = RES.o$sequence[m]

head(RES.c, 20)

save(RES.c, file = paste0(IO$output_Rdata, "RES_2.Rdata"))


```




```{r impact_of_noise  visualizing results of conservative HMM, cache = TRUE}

g = ggplot(RES.c, aes(x = ovu.diff, fill = C))
g + geom_histogram(bins = 50) + xlim(-2.5,2.5) + facet_grid(C~.)


g = ggplot(RES.c, aes(x = dev.sym, fill = C))
g + geom_histogram(bins = 50)  + facet_grid(C~.)


g = ggplot(RES.c, aes(x = abs(ovu.diff), y = ovu.sd.no.noise, col = C))
g  + geom_abline(slope = 1, intercept = 0)+ geom_point(size = 0.5) + facet_wrap(C ~.)


```






```{r impact_of_noise  checking the state sequence differences for conservative hmm, cache = TRUE}

STD.c = data.frame()

#plot(c(0,50),c(0,nrow(RES)), type = "n")
for(i in 1:nrow(RES.c)){
  seq = RES.c$sequence[i]
  seq = unlist(strsplit(as.character(seq), split = " "))
  seq.num = match(seq, hmm_par$cycle.states)
  
  seq.o = RES.c$sequence.no.noise[i]
  seq.o = unlist(strsplit(as.character(seq.o), split = " "))
  seq.o.num = match(seq.o, hmm_par$cycle.states)
  
  #points(rep(i,length(seq)), col = cycle.states.colors[seq.num], pch = "-")
  
  dist.seq = sum(seq.num != seq.o.num)
  n.back.trans = sum(diff(seq.num) == -1)
  period.duration = sum(seq.num %in% c(1,2))
  period.duration.o = sum(seq.o.num %in% c(1,2))
  
  line = data.frame(cycle_id = RES.c$cycle_id[i],
                    y = RES.c$y[i],
                    C = RES.c$C[i],
                    dist.seq = dist.seq,
                    cycle_length = length(seq.num),
                    n.back.trans = n.back.trans,
                    period.duration = period.duration,
                    period.duration.change = period.duration-period.duration.o)
  STD.c = rbind(STD.c, line)
  
}

STD.c$dist.seq.f = STD.c$dist.seq/STD.c$cycle_length * 100


g = ggplot(STD.c, aes(x = dist.seq.f, fill = factor(C)))
g + geom_histogram() + facet_grid(C~.)


g = ggplot(STD.c, aes(x = n.back.trans, fill = factor(C)))
g + geom_histogram() + facet_grid(C ~.)





```




## Conclusion

It actually changes very little in this case as well.
The constrain on the end state and the emission probabilities seem to make the ovulation detection quite robust.


What would be interesting, would be to modify the emission probabilities and check if they have a stronger effect on the ovulation detection.




# Noise on the emission probabilities



```{r impact_of_noise running modified HMM with changing emission probabilities, cache = TRUE}

Y = 9

RES.e = data.frame()

tic()
for (cycle_id in selected_cycle_ids) {
  #cat(cycle_id, '\n')
  j = which(cycledays$cycle_id == cycle_id)
  cycletable = cycledays[j, ]
  
  for(y in 1:Y){
    #cat("\t", y, "\n")
    
    #noise = -log10(y/Y*10)+1
    noise = y/Y * 3
    
    res = try(most_likely_day_of_ovulation_hmm(cycletable = cycletable,
                                               debug = FALSE, no.print = TRUE,
                                               random_emission_probabilities = TRUE,
                                               noise = noise),
              silent = TRUE)
    
    if(typeof(res)!= 'character'){
      #cat("\t\there\n")
      res$cycle_id = cycle_id
      line = data.frame(
        cycle_id = cycle_id, 
        y = y,
        noise = noise,
        dev.sym = res$dev.sym,
        diff.em = res$diff.em,
        ovu = res$ovu, 
        ovu.sd = res$ovu.sd,
        ovu.viterbi = res$ovu.viterbi, 
        confidence = res$confidence,
        sequence = paste(res$obs.viterbi, collapse = " "))
      RES.e = rbind(RES.e, line)
    }
    
  }
  
}
toc()

m = match(RES.e$cycle_id, RES.o$cycle_id)
RES.e$ovu.no.noise = RES.o$ovu[m]
RES.e$ovu.diff = RES.e$ovu.no.noise - RES.e$ovu
RES.e$ovu.sd.no.noise = RES.o$ovu.sd[m]
RES.e$sequence.no.noise = RES.o$sequence[m]
RES.e$noise.r = round(RES.e$noise, digits = 2)

head(RES.e, 20)



save(RES.e, file = paste0(IO$output_Rdata, "RES_3.Rdata"))

```



```{r impact_of_noise visualizing results of random emission HMM , cache = TRUE}


g.em = ggplot(RES.e, aes(x = ovu.diff, fill = noise))
g.em = g.em + geom_histogram(bins = 100) + 
  facet_grid(noise.r ~.) + 
  scale_fill_gradient(low = "gray30", high = "green3")+
  guides(fill=FALSE)+
  scale_y_continuous(breaks = seq(0,100,by =30))+
  theme(strip.text.y = element_text(angle = 0))+
  #ggtitle("difference in ovulation estimation") + 
  xlab("Diff. in ovu. est. (no noise - noise) [days]") + xlim(-xlim.max, xlim.max) 

g.em


g.em.scatter = ggplot(RES.e, aes(x = abs(ovu.diff), y = ovu.sd.no.noise, col = noise.r))
g.em.scatter = g.em.scatter  + 
  geom_abline(slope = 1, intercept = 0, col = "gray")+ 
  geom_point(size = 0.5) + 
  scale_color_gradient(low = "gray30", high = "green3")+
  facet_wrap(noise.r ~., ncol = 5)+
  xlab("Abs. diff. in ovu. est. (no noise - noise) [days]")+
  ylab("Uncertainty on ovulation est. (no noise)")+
  guides(col=FALSE) + xlim(0,xlim.max) 

g.em.scatter




g.em.sd = ggplot(RES.e, aes(x = ovu.sd, fill = noise))
g.em.sd = g.em.sd + geom_histogram(bins = 100) + 
  facet_grid(noise.r ~.) + 
  scale_fill_gradient(low = "gray30", high = "green3")+
  guides(fill=FALSE)+
  scale_y_continuous(breaks = seq(0,100,by =30))+
  theme(strip.text.y = element_text(angle = 0))+
  #ggtitle("difference in ovulation estimation") + 
  xlab("Ovu. est. uncertainty [days]") 

g.em.sd



g.em.diff = ggplot(RES.e, aes(x = diff.em, fill = noise))
g.em.diff = g.em.diff + geom_histogram(bins = 100) + 
  facet_grid(noise.r ~.) + 
  scale_fill_gradient(low = "gray30", high = "green3")+
  guides(fill=FALSE)+
  scale_y_continuous(breaks = seq(0,100,by =30))+
  theme(strip.text.y = element_text(angle = 0))+
  #ggtitle("difference in ovulation estimation") + 
  xlab("Ovu. est. uncertainty [days]") 

g.em.diff



```




